 Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021.

### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization: Fourth commit, includes updated README, and output documents such as paper 1.qmd and paper 1.pdf

---
title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021. "
author: Sofia Sellitto
date: February 3, 2023
date-format: long
format: 
  pdf: 
    documentclass: article 
    geometry: margin = 1in
    abstract: In the city of Toronto, citizens have become accustomed to fearing for their safety as crime appears to have increased in recent years. In this paper, I analyze the number of victims of crimes against the person in the City of Toronto from 2014 - 2021 to investigate if the number of Torontonians victimized by crime has increased over time and if sex differences are apparent across various categories of crime. By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time, sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female. These results have implications for how Torontonians view their safety and consider how their sex impacts the possibility of crimes they fall victim to. 
toc: true
toc-depth: 4
theme: cosmo
bibliography: bibliography.bib
---

Code and data supporting this analysis is available at: LINK

@@

Prior to the 2006 Canadian federal election, Stephen Harper, who would eventually be elected Prime Minister until 2016, led a conservative campaign based on crime prevention (cite SH book). The conservative party asserted that Canada was facing a period of increased crime and assured protection to Canadians by way of tougher sentencing/ parole, additional police funding, and investments in victim services, among a myriad of other things [@book]. In his 2011 campaign, Harper suggested that judicial opinions favoured the rights of criminals instead of Canadian citizens [@book]. Harper's claims of increased crime and strides for harsher punishment sparked debates of fear-mongering and inducing an era of incorrectly generated moral panic. According to statistics Canada, the total rate of crime had declined over 20 years since its peak in 1991 [@book]. This was in direct contradiction to Harper's claims.

@@

In this paper, I utilize Open Data Toronto [@opendatatoronto] to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. This is to investigate how the number of victims of crime in Toronto has changed over time, and if it has increased  over time, which categories of crimes against the person are most prevalent and if sex differences are apparent across different forms of crime. I will commence with a discussion of my data source and its features. I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violence" and "other", across 2014-2019. Next, I will move towards an analysis of how the reported number of victims has changed according to the category of crime, over time. I will conclude my analysis with an investigation of how the number of victimization reports for each category of crime, differs according to the gender of the victim. The paper will culminate with a discussion of my findings, limitations and future research possibilities.

@@

The data source utilized for the paper was taken from the City of Toronto's open data portal [@opendatatoronto]. The data was collected from the Toronto Police Service Annual Statistical Report (ASR) which conducts overviews of police-related statistics. For this paper specifically, the dataset is comprised of all identified victims of crimes against the person (victims of crimes) from 2014-2021.

@@

\
This paper will follow an analysis of the following variables: the reported year of the crime, the crime category subtype, the number of identified victims and their respective sex. I have chosen to exclude the assault subtype of the crime as it included numerous missing values. I would have been forced to remove this data, thus negatively impacting my analysis. Because my analysis is focused on victims of crime over time and potential gender differences, I have decided to withdraw from using data for age group or cohort.

\
Included within the crime category subtype are (1) assault, (2) robbery, (2) sexual violation and (4) other. To further discuss these categorizations, in Canada, assault is classified as intentionally applying force to an individual indirectly or directly without consent. This can then include assault to cause bodily harm, with a deadly weapon, aggravated assault or attempted murder. In Canada, robbery is defined as stealing for extortion of the article stolen and can include the use of violent threats, personal violence or with a weapon. In Ontario, sexual violence describes any physical or psychological violence conducted or targeted in a sexual means. Unfortunately, no information was provided by the City of Toronto nor the Toronto Police Service Annual Statistical Report (ASR) on the nature of the "other" category. We are forced to assume that this may include break and enter, homicide, domestic violence or various forms of crimes against the person.

@@

I initiated my analysis by cleaning and sorting my data. I first selected the variables of interest, checked for missing values, cross-tabulated to understand the pattern of unidentified data and dropped unknown variables. Included in the sex variable was "U" for unknown, which unfortunately left me forced to remove it from my dataset. This is a limitation of my analysis that I will further examine in my discussion.

My investigation began with an overview of the average number of people victimized by crime in Toronto across 2014-2021 based on the sub-category of crime. My intent with this is to help visualize the numerical average of Torontonians who were the victims of each sub-category of crime across 2014 - 2021. I visualized this in a table (see table 1) to create categorical organization from the highest to lowest average victims of crime.

@@

knitr::kable(mean_data,
             "simple", 
             caption = "Average Number of victims of crimes in Toronto across 2014                         -2021, grouped by category of crime", 
             col.names=c("Category of Crime", "Average Number of Victims"), 
             align=c('l', 'c'))

@@

knitr::kable(head(sex_data), 
             "simple", 
             caption = "The first 6 rows of the data frame used to represent the                           gender of the number of victims according to category of                           crime in Toronto Across 2014 - 2021", col.names=c("Sub                             -Category of Crime", "Sex", "Year", "Number of Victims"), 
             align=c('l', 'c', 'c'))

@@

As per my primary objective for this paper, I wish to understand how the number of victims of crime in Toronto has changed over time and to analyze if victimization numbers are increasing over time. Because of this, I created a line graph that showcases the number of victims according to each subcategory of crime over 2014 -2019 (See Figure 2). To do this, I called on the data frame visualized above (See Table 2). This visualization allows for comparison across the different subcategories of crimes while simultaneously presenting a rich depiction of how the number of victims has changed over time, the year they peaked and at their lowest. Using the data previewed in Table 1, I was able to discriminate the number of total victims for the peak years of each sub-category of crime.

@@

By drawing your attention to Figure 1, it is apparent from 2014-2021, the most prevalent crime that Torontonians were victims of was "Assault". From 2014 - 2018, the number of annual victims gradually increased until it peaked in 2019, with 18819 total victims. The number of victims of assault then began to decrease following 2019, where it spiked in 2021 but remained below 17500 victims.

@@

"Robbery" had the third most victims of crime from 2014-2021. The peak number of victims for this variable is not very distinct, but saw its highest in 2017 with 3686 victims that year. From 2017 onward the number of victims of robberies steadily decreased.

"Sexual Violence" had the least number of victims of crime from 2014-2021. However, contrary to the others, sexual violence saw its peak in 2021 with 2568 victims. From 2014 onward, the number of victims gradually increased until it peaked in 2021.

In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "Sexual Violence". Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively. Using the data set from Table 2, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.

@@

\
Lastly, "Sexual Violence", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violence than men. This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021 being female, and 334 being male. This means a difference of 147.9%.

@@

In this paper, I have utilized The City of Torontos open data portal, supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014-2021. The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed and if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent across the categories of crime. By utilizing the Coding language R (cite) I was able to import, clean and asses the raw data made available by the City. In doing so I discovered that for the sub-category of crimes, assault, robbery, and other, the number of Torontonian victims peaked in the years 2019, 2017, and 2015 respectively and decreased until the end of the data collection date (2021). The most prevalent crime that Torontonians were victims of was assault. Contrarily, sexual violence saw its peak in 2021, meaning it has not decreased over time. In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violence had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%. Robbery maintains the second largest difference between male and female victims, skewing towards men with a difference of 101.35%. "Other", had the third largest difference, with more female victims than male. Assault had the least large sex difference, with slightly more male victims than feel with a difference of 15.53%.

@@

Because of the nature of the data set, my analysis is limited in many regards. To begin with the source, Torontos Police Service Annual Statistical Report (ASR), it is possible that data was misrepresented, incorrectly input, missing, duplicated or potentially falsified. While it is important to avoid speculation about these statistics' validity, I believe it is necessary to highlight possibilities for error. Another vital point to acknowledge is the number of individuals who may have fallen victim to a crime against the person and failed to report it or those who intended to file a report but were dismissed. The latter is an extremely apparent phenomenon for women who attempt to report sexual violence. While this data represents the number of individuals who have fallen victim to crimes against the person across 2014-2021, it cannot account for unfiled and dismissed reports. Meaning that the number of victims could potentially be far higher than portrayed in the analysis. Regarding the data set itself, removing missing and unknown data will impact the analysis in various regards. More specifically, regarding the variable sex variable including "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination. Removing this data could have prevented insightful analysis of non-binary and Agender individuals.

>>>>

---
title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021."
author: Sofia Sellitto
date: February 3, 2023
date-format: long
format:
  pdf: 
    documentclass: article 
    geometry: margin = 1in
    abstract: In the city of Toronto, citizens have become accustomed to fearing for their safety as crime appears to have increased in recent years. In this paper, I analyze the number of victims of crimes against the person in the City of Toronto from 2014 - 2021 to investigate if the number of Torontonians victimized by crime has increased over time and if sex differences are apparent across various categories of crime. By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time. Sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female. These results have implications for how Torontonians view their safety and consider how sex impacts their predisposition to certain crimes. 
toc: true
toc-depth: 4
theme: cosmo
bibliography: bibliography.bib
---

^[Code and data supporting this analysis is available at: LINK]

@@

Prior to the 2006 Canadian federal election, Stephen Harper, who would eventually be elected Prime Minister until 2016, led a conservative campaign based on crime prevention [@book]. The conservative party asserted that Canada was facing a period of increased crime and assured protection to Canadians by way of tougher sentencing/ parole, additional police funding, and investments in victim services, among a myriad of other things [@book]. In his 2011 campaign, Harper suggested that judicial opinions favoured the rights of criminals instead of Canadian citizens [@book]. Harper's claims of increased crime and strides for harsher punishment sparked debates of fear-mongering and inducing an era of incorrectly generated moral panic. According to statistics Canada, the total rate of crime had declined over 20 years since its peak in 1991 [@book]. This was in direct contradiction to Harper's claims.

@@

In this paper, I utilize Open Data Toronto [@opendatatoronto] to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. This is to investigate how the number of victims of crime in Toronto has changed over time, if it is increasing, which categories of crimes are most prevalent and if sex differences are apparent across different forms of crime. I will commence with a discussion of my data source and its features. I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violence" and "other", across 2014-2019. Next, I will move towards an analysis of how the number of victims has changed according to the category of crime, over time. Following, I will conduct an investigation of how the number of victimization reports for each category of crime, differs according to the sex of the victim. The paper will culminate with a discussion of my findings, limitations and future research possibilities.

@@

The data source utilized for this paper was taken from the City of Toronto's open data portal [@opendatatoronto]. The data was collected from the Toronto Police Service Annual Statistical Report (ASR) which conducts overviews of police-related statistics. For this paper specifically, the dataset is comprised of all identified victims of crimes against the person (victims of crimes) from 2014-2021.

@@

\
This paper will follow an analysis of the following variables: the reported year of the crime, the crime category subtype, the number of identified victims and their respective sex. I have chosen to exclude the assault subtype of the crime as it included numerous missing values. I would have been forced to remove this data, thus negatively impacting my analysis. Because I am focused on victims of crime over time and potential gender differences, I have withdrawn from using data on age group or cohort.

\
Included within the crime category subtype are (1) assault, (2) robbery, (3) sexual violation and (4) other. To further discuss these categorizations, in Canada, assault is classified as intentionally applying force to an individual indirectly or directly without consent. This can then include assault to cause bodily harm, with a deadly weapon, aggravated assault or attempted murder. In Canada, robbery is defined as stealing for extortion of the article stolen and can include the use of violent threats, personal violence or with a weapon. In Ontario, sexual violence describes any physical or psychological violence conducted or targeted in a sexual means. Unfortunately, no information was provided by the City of Toronto nor the Toronto Police Service Annual Statistical Report (ASR) on the nature of the "other" category. We are forced to assume that this may include break and enter, homicide, domestic violence or various forms of crimes against the person.

@@

I initiated my analysis by cleaning and sorting my data. I first selected the variables of interest, checked for missing values, cross-tabulated to understand the pattern of unidentified data and dropped unknown variables. Included in the sex variable was "U" for unknown, which I was unfortunatelyforced to remove it from my dataset. This is a limitation of my analysis that will be further examined in my discussion.

My investigation began with an overview of the average number of people victimized by crime in Toronto across 2014-2021 based on the sub-category of crime. I visualized this data to create categorical organization from the highest to lowest average victims of crime, see table 1.

@@

knitr::kable(mean_data,
             "simple", 
             caption = "Average number of victims of crimes in Toronto across 2014                         -2021, grouped by category of crime", 
             col.names=c("Category of Crime", "Average Number of Victims"), 
             align=c('l', 'c'))

```
From this table, it is apparent that assault has the highest average number of victims total spanning 2014 - 2021.

@@

knitr::kable(head(sex_data), 
             "simple", 
             caption = "The first 6 rows of the data frame used to represent the                           sex of the number of victims according to category of crime in Toronto Across 2014 - 2021", col.names=c("Sub-Category of Crime", "Sex", "Year", "Number of Victims"), 
             align=c('l', 'c', 'c'))

@@

As per my primary objective for this paper, I wish to understand how the number of victims of crime in Toronto has changed over time and to analyze if victimization numbers are increasing over time. Because of this, I created a line graph that showcases the number of victims according to each subcategory of crime over 2014 - 2019 (See Figure 2). To do this, I called on the data frame visualized above (See Table 2). This visualization allows for comparison across the different subcategories of crimes while simultaneously presenting a rich depiction of how the number of victims has changed over time, the year they peaked and at their lowest. Using the data previewed in Table 2, I was able to discriminate the number of total victims for the peak years of each sub-category of crime.

@@

By drawing your attention to Figure 1, it is apparent from 2014 - 2021, the most prevalent crime that Torontonians were victims of was "Assault". From 2014 - 2018, the number of annual victims gradually increased until it peaked in 2019, with 18819 total victims. The number of victims of assault then began to decrease following 2019, where it spiked in 2021 but remained below 17500 victims.

@@

"Robbery" had the third most victims of crime from 2014 - 2021. The peak number of victims for this variable is not very distinct, but saw its highest in 2017 with 3686 victims that year. From 2017 onward the number of victims of robberies steadily decreased.

"Sexual Violence" had the least number of victims of crime from 2014 - 2021. However, contrary to the others, sexual violence saw its peak in 2021 with 2568 victims. From 2014 onward, the number of victims gradually increased until it peaked in 2021.

In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "Sexual Violence". Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively. Using the data set from Table 3, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.

@@

\
Lastly, "Sexual Violence", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violence than men. This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021, being female, and 334 being male. This means a difference of 147.9%.

@@

In this paper, I have utilized The City of Torontos open data portal [@opendatatoronto], supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed, if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent. By utilizing the Coding language R [@R] I was able to import, clean and asses the raw data made available by the City. In doing so I discovered that for the sub-category of crimes, assault, robbery, and other, the number of Torontonian victims peaked in the years 2019, 2017, and 2015 respectively and decreased until the end of the data collection date (2021). The most prevalent crime that Torontonians were victims of was assault, maintaining the highest cumulative average across 2014-2021. Contrarily, sexual violence saw its peak in 2021, meaning it has not decreased over time. In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violence had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%. Robbery maintains the second largest difference between male and female victims, skewing towards men with a difference of 101.35%. "Other", had the third largest difference, with more female victims than male. Assault had the least large sex difference, with slightly more male victims than feel with a difference of 15.53%.

@@

Because of the nature of the data set, my analysis is limited in many regards. To begin with the source, Torontos Police Service Annual Statistical Report (ASR), it is possible that data was misrepresented, incorrectly input, missing, duplicated or potentially falsified. While it is important to avoid speculation about these statistics' validity, I believe it is necessary to highlight possibilities for error. Another vital point to acknowledge is the number of individuals who may have fallen victim to a crime against the person and failed to report it or those who intended to file a report but were dismissed. The latter is an extremely apparent phenomenon for women who attempt to report sexual violence. While this data represents the number of individuals who have fallen victim to crimes against the person across 2014 - 2021, it cannot account for unfilled and dismissed reports. Meaning that the number of victims could potentially be far higher than portrayed in the analysis. Regarding the data set itself, removing missing and unknown data will impact the analysis in various regards. More specifically, concerning the variable sex including the categories "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination. Removing this data could have prevented insightful analysis of non-binary and Agender individuals.
